---
title: "Thalamus Analysis"
author: "FG"
date: "5 3 2021"
output: 
  bookdown::html_document2:
    code_folding: hide
    toc: yes
    toc_depth: 5
    highlight: tango
    theme: paper
    fig_width: 20
    fig_height: 20
    fig_caption: yes
    number_sections: yes
    keep_md: yes
    df_print: default
---

```{r library, results='hide', message=FALSE, warning=FALSE} 
library(htmlTable)
library(rmarkdown)
library(Hmisc)
library(dplyr)
library(psych)
library(pastecs)
library(tidyr)
library(tidyverse)
library(ggplot2) # for data visualization (plotting)
library(car) # for statistical models (ANCOVA with type III errors), dummy coding
library(MatchIt)
library(ggforce) # extends functionality of ggplot2 (e.g. spreads boxplot facets (facet_wrap) over multiple pages)
library(cowplot)
library(ggsci) # alternative color tool/palettes for plotting (scientific journal style)
library(ggpubr) # extends functionality of ggplot2 (e.g. editing options for publication)
library(ggsignif)
library(kableExtra) # for creating kables (customized tables in RMarkdown output with styling options)
library(bookdown) # support for numbering figure and table captions
library(rstatix) 
library(MASS) # for robust linear model
library(broom) # tidy test output, e.g. of GLM
library(emmeans) # for estimated marginal means (EMM) (covariate-adjusted means in ANCOVA)
library(sfsmisc)
library(scales)
library(sjstats)
library(sqldf)
```

# Vorbereitungen
*1. set path to taurus mountpoint*
```{r filepaths, results='hide', message=FALSE, warning=FALSE}
taurus_mountpoint <- first(Filter(dir.exists, 
                               c("/Applications/freesurfer/7.1.1/subjects/scratch/igis/processing/nipype/smri_7",
                                 "/scratch/igis/processing/nipype/smri_7")))
```

*2. getting subfield volumes*
```{r fssubdata, results='hide', message=FALSE, warning=FALSE}
get_subfield_volumes <- function(){
  
  cache_file <- "cache/subfield_volumes.csv"
  
  if (!file.exists(cache_file)){  
    filenames <- Sys.glob(file.path(taurus_mountpoint, "*/mri/*.txt")) # file paths to mounted FreeSurfer output - .txt files of all FreeSurfer-processed subjects via wildcard expansion
    
    tibble(file=filenames) %>%
      filter(str_detect(file, "(.*Thalamic.*)")) %>%
      mutate(content=map(file, function(x) { # content refers to nested "region" and "measure" data
        read_delim(x, delim=" ", 
                   col_names=c("region" , "measure"),
                   col_types=c(region=col_character(),
                               measure=col_double()))
      })) %>%
      unnest(content) %>%
      extract(file, c("storage_name", "file_type"), ".*/(64-[0-9][0-9]-[0-9][0-9][0-9]-[0-9]_[^/]*)/.*/([^/]*)", remove=FALSE) %>%
      extract(file_type, c("hemi"), "([rl]h).*", remove=FALSE ) %>%
      mutate(hemi = ifelse(startsWith(region, "Right"),"rh", hemi)) %>% # only for thalamic nuclei ("Left/Right" included in filenames)
      mutate(hemi = ifelse(startsWith(region, "Left"),"lh", hemi)) %>%
      mutate(region = str_remove(region, "(Right-)|(Left-)")) %>%
      mutate(analysis = dplyr::recode(file_type,
                                      "ThalamicNuclei.v12.T1.volumes.txt" = "thalamus",)) %>%
      select(-file) %>%
      write_csv(cache_file)
  }
  
  return(read_csv(cache_file, col_types = cols()))  
}
```

*3. getting aseg volumes*
```{r fsbraindata1, results='hide', message=FALSE, warning=FALSE}
get_aseg_volumes <- function(){
  
  cache_file <- "cache/aseg_volumes.csv"
  
  if (!file.exists(cache_file)){
      filenames_aseg_stats <- Sys.glob(file.path(taurus_mountpoint, "*/stats/aseg.stats"))
      
      tibble(filenames_aseg_stats) %>%
        mutate(storage_name = str_extract(filenames_aseg_stats, "64-[0-9][0-9]-[0-9][0-9][0-9]-[0-9]_[^/]*")) %>%
        mutate(content = map(filenames_aseg_stats, function(x) {
          read_file(x)
        })) %>%
        unnest(content) %>%
        extract(content, "e_tiv", "EstimatedTotalIntraCranialVol.*,\ ?(\\d+\\.*\\d*),", remove=F, convert=T) %>%
        extract(content, "brain_seg_not_vent", "BrainSegNotVent.*,\ ?(\\d+\\.*\\d*),", remove=F, convert=T) %>%
        extract(content, "supra_tentorial_not_vent", "SupraTentorialNotVent.*,\ ?(\\d+\\.*\\d*),", remove=F, convert=T) %>%
        extract(content, "total_gray", "TotalGray.*,\ ?(\\d+\\.*\\d*),", remove=F, convert=T) %>%
        extract(content, "subcort_gray", "SubCortGray.*,\ ?(\\d+\\.*\\d*),", remove=F, convert=T) %>%
        select(-content, -filenames_aseg_stats) %>%
        write_csv(cache_file)
    }
  
    return(read_csv(cache_file, col_types = cols()))  
}
```

*4. getting CNR*
```{r getcnr, results='hide', message=FALSE, warning=FALSE}
get_cnr <- function(){
  
  cache_file <- "cache/cnr.csv"
  
  if (!file.exists(cache_file)){  

    filenames <- Sys.glob(file.path(taurus_mountpoint, "*/stats/norm_CNR.txt"))
    tibble(filenames) %>%
      mutate(storage_name = str_extract(filenames, "64-[0-9][0-9]-[0-9][0-9][0-9]-[0-9]_[^/]*")) %>%
      mutate(content = map(filenames, function(x) {
        read_lines(x, skip=6) # start reading at line 7 = last line of the file which contains the cnr_total measure
      })) %>%
      unnest(content) %>%
      extract(content, "cnr_total", "total CNR = (.*)$", convert=T) %>%
      select(-filenames) %>%
      write_csv(cache_file)
  }

  return(read_csv(cache_file, col_types = cols()))  
}
```

*5. getting SNR*
```{r getsnr, results='hide', message=FALSE, warning=FALSE}
get_snr <- function(){
  
  cache_file <- "cache/snr.csv"
  
  if (!file.exists(cache_file)){  
    
    filenames <- Sys.glob(file.path(taurus_mountpoint, "*/stats/wmsnr.e3.dat"))
    tibble(filenames) %>%
      mutate(content = map(filenames, function(x) {
        read.table(x, sep="", 
                   col.names=c("storage_name", 
                               "snr", 
                               "var3",
                               "var4",
                               "var5",
                               "var6")) })) %>%
      unnest(content) %>%
      select(storage_name, snr) %>%
      write_csv(cache_file)
  }
  
  return(read_csv(cache_file, col_types = cols()))  
}
```


*6. getting redcap data*
```{r redcapimport, results='hide', message=FALSE, warning=FALSE}

get_redcap <-function(){
  read_csv("data/redcap_do_not_delete/redcap.csv", col_types = cols())
}  %>%
  filter(!str_detect(storage_name, "atd" )) %>% # exclude atd study arm
  filter(!str_detect(storage_name, "^25" ))
```

*7. applying exclusion criteria before matching*

# 1. exclusion based on redcap data
exclude psychiatric comorbidities (in HC), medication (except SSRIs in T1)
```{r} 
redcap.medical <- get_redcap() %>%
  filter(point_of_research == "HCW") %>%
  filter(psychiatric_disorders == "nein") %>%
  dplyr::select(participant_id,point_of_research,storage_name,current_antidepressants_preparation,type_of_current_medication___1, type_of_current_medication___2, type_of_current_medication___3, type_of_current_medication___4, type_of_current_medication___5) %>%
    rename(current_antidepressants = type_of_current_medication___1)%>%
  rename(current_neuroleptic = type_of_current_medication___2)%>%
  rename(current_stimulantien_MPH = type_of_current_medication___3)%>%
  rename(current_mood_stabilizers = type_of_current_medication___4)%>%
  rename(current_benzodiazepine = type_of_current_medication___5) %>%
  group_by(participant_id) %>%
  slice_head(n = 1) %>%
  ungroup()

get_redcap() %>%
  filter(point_of_research == "T1") %>% 
  dplyr::select(participant_id, storage_name,point_of_research, current_antidepressants_preparation, type_of_current_medication___1, type_of_current_medication___2, type_of_current_medication___3, type_of_current_medication___4, type_of_current_medication___5) %>%
  rename(current_antidepressants = type_of_current_medication___1)%>%
  rename(current_neuroleptic = type_of_current_medication___2)%>%
  rename(current_stimulantien_MPH = type_of_current_medication___3)%>%
  rename(current_mood_stabilizers = type_of_current_medication___4)%>%
  rename(current_benzodiazepine = type_of_current_medication___5) %>%
  filter(!current_neuroleptic == "Checked") %>%
  group_by(participant_id) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  full_join(redcap.medical) -> redcap.medical 
```

# 2. sMRI data 
this is a three-step exclusion, startin with 1) exclude all subjects that are not labeled as "Einschluss" in variable "mriqc_smri_thal_summary", 2) exclude all subjects that are not labelled as 0, 1 in variable "mriqc_smri_summary", 3) exclude outliers via variable "outliersum = n_outliers_volume_2.698sd + n_outliers_symmetry_2.698sd". This variable is generated as the sum from volume and symmetry outliers that are imported via file "thal_subcortical_QC.csv" in the data folder.

```{r}

```

# 3. BMI(-SDS) exclusion
in this step we exclude high and low BMI for HC: >28 and <17.5 (>= 18 years of age) are excluded. If participants are ouside this range, age has to be checked. If they are unter 18, get BMI-SDS for this age. If at least 18, the subject must be excluded. Same is for T1: subjects must have a max BMI of 17.5, if more and younger than 18: check BMI-SDS. John already checked affected subjects (see code below).

```{r}

```

*8. get matched sample*

```{r}

```







# get matched sample to work with
```{r}
sample <- T1_exclusion %>% 
  rbind(HC_exclusion) %>%
  filter(mriqc_smri_thal_summary == "Einschluss") %>%
  filter(type_of_current_medication___2 == "Unchecked")%>% # exclude neuroleptika
  mutate(point_of_research_n = dplyr::recode(point_of_research,
                                    "HCW" = 0,
                                    "T1" = 1)) %>% 
  group_by(participant_id) %>%
  slice(1) %>%
  ungroup() %>%
  matchit(point_of_research_n ~ age_at_date_of_research, data = ., method = "optimal", ratio = 1) %>%
  match.data()

stopifnot(nrow(sample) == 340)
```



### 64-10-372-1, 64-10-200-1, 64-10-429-1
```{r}
sample_t1_disorders <- sample %>%
  filter(point_of_research %in% "T1") %>%
  filter(psychiatric_disorders == "ja") %>%
  dplyr::select(participant_id,point_of_research,psychiatric_disorders,
                                    depressive_disorders, 
                                    obsessivecompulsive_disorders, 
                                    anxiety_disorders, 
                                    addiction, 
                                    ptsd, 
                                    tic_tourette,
                                    conduct_disorder,
                                    defiant_behaviour,
                                    adaption_disorder,
                                    psychosis,
                                    attention_deficit_hyperactivity_disorder,
                                    personality_disorders,
                                    developmental_disorders,
                                    other_psychiatric_disorders,
                                    spec_other_psychiatric_disorders) %>%
  mutate_all(funs(str_replace(., "unbekannt", NA_character_)))%>%
  mutate(n_current_depressive_disorders = sum(str_count(depressive_disorders, "ja"))) %>%
  mutate(n_current_obsessivecompulsive_disorders = sum(str_count(obsessivecompulsive_disorders, "ja")))%>%
  mutate(n_current_anxiety_disorders = sum(str_count(anxiety_disorders, "ja"))) %>%
  mutate(n_current_addiction = sum(str_count(addiction, "ja"))) %>%
  mutate(n_current_ptsde = sum(str_count(ptsd, "ja"))) %>%
  mutate(n_current_conduct_disorder = sum(str_count(conduct_disorder, "ja"))) %>%
  mutate(n_current_defiant_behaviour = sum(str_count(defiant_behaviour, "ja"))) %>%
  mutate(n_current_adaption_disorder = sum(str_count(adaption_disorder, "ja"))) %>%
  mutate(n_current_psychosis = sum(str_count(psychosis, "ja"))) %>%
  mutate(n_current_adhd = sum(str_count(attention_deficit_hyperactivity_disorder, "ja"))) %>%
  mutate(n_current_personality_disorders = sum(str_count(personality_disorders, "ja"))) %>%
  mutate(n_current_developmental_disorders = sum(str_count(developmental_disorders, "ja"))) %>%
  mutate(n_current_other = sum(str_count(other_psychiatric_disorders, "ja"))) %>%
  mutate(n_current_spec_other = sum(str_count(spec_other_psychiatric_disorders, "ja")))
  
```




*9. getting FS data for sample*
```{r getGLMdata, results="asis", message=FALSE, warning=FALSE}
glmdata <- sample %>% 
  dplyr::select(storage_name, point_of_research, age_at_date_of_research) %>%
  left_join(get_aseg_volumes()) %>%
  left_join(get_cnr()) %>%
  left_join(get_snr()) %>%
  left_join(get_subfield_volumes() %>% filter(analysis == "thalamus")) %>%
  mutate(region_recoded = dplyr::recode(region,
                                       'AV' = "Anteroventral",
                                       'CL' = "Central lateral",
                                       'CM' = "Centromedian",
                                       'CeM' = "Central medial",
                                       'L-Sg' = "Limitans (suprageniculate)",
                                       'LD' = "Laterodorsal",
                                       'LGN' = "Lateral geniculate",
                                       'LP' = "Lateral posterior",
                                       'MDl' = "Mediodorsal lateral parvocellular",
                                       'MDm' = "Mediodorsal medial magnocellular",
                                       'MGN' = "Medial Geniculate",
                                       'MV(Re)' = "Reuniens (medial ventral)",
                                       'Pc' = "Paracentral",
                                       'Pf' = "Parafascicular",
                                       'Pt' = "Paratenial",
                                       'PuA' = "Pulvinar anterior",
                                       'PuI' = "Pulvinar inferior",
                                       'PuL' = "Pulvinar lateral",
                                       'PuM' = "Pulvinar medial",
                                       'VA' = "Ventral anterior",
                                       'VAmc' = "Ventral anterior magnocellular",
                                       'VLa' = "Ventral lateral anterior",
                                       'VLp' = "Ventral lateral posterior",
                                       'VM' = "Ventromedial",
                                       'VPL' = "Ventral posterolateral",
                                       'Whole_thalamus' = "Whole Thalamus")) %>%
  unite("hemi_region", c(hemi, region), sep = "_", remove = FALSE, na.rm = FALSE) %>%
  unite("labeller", c(region_recoded, hemi), sep= " ", remove = FALSE, na.rm = FALSE) %>% # for output tables (labels without underscores)
  mutate(hemi_region = factor(hemi_region, levels=c('lh_Whole_thalamus',
                                                   'rh_Whole_thalamus',
                                                   'lh_AV','rh_AV',
                                                   'lh_LD','rh_LD',
                                                   'lh_LP','rh_LP',
                                                   'lh_VA','rh_VA',
                                                   'lh_VAmc','rh_VAmc',
                                                   'lh_VLa','rh_VLa',
                                                   'lh_VLp','rh_VLp',
                                                   'lh_VPL','rh_VPL',
                                                   'lh_VM','rh_VM',
                                                   'lh_CeM','rh_CeM',
                                                   'lh_CL','rh_CL',
                                                   'lh_Pc','rh_Pc',
                                                   'lh_CM','rh_CM',
                                                   'lh_Pf','rh_Pf',
                                                   'lh_Pt','rh_Pt',
                                                   'lh_MV(Re)','rh_MV(Re)',
                                                   'lh_MDm','rh_MDm',
                                                   'lh_MDl','rh_MDl',
                                                   'lh_LGN','rh_LGN',
                                                   'lh_MGN','rh_MGN',
                                                   'lh_L-Sg','rh_L-Sg',
                                                   'lh_PuA','rh_PuA',
                                                   'lh_PuM','rh_PuM',
                                                   'lh_PuL','rh_PuL',
                                                   'lh_PuI','rh_PuI'))) %>%
    mutate(group = dplyr::recode(region, 'AV' = "anterior", 
                           'CL' = "intralaminar",
                           'CM' = "intralaminar",
                           'CeM' = "intralaminar",
                           'L-Sg' = "posterior",
                           'LD' = "lateral",
                           'LGN' = "posterior",
                           'LP' = "lateral",
                           'MDl' = "medial",
                           'MDm' = "medial",
                           'MGN' = "posterior",
                           'MV(Re)' = "medial",
                           'Pc' = "intralaminar",
                           'Pf' = "intralaminar",
                           'Pt' = "medial",
                           'PuA' = "posterior",
                           'PuI' = "posterior",
                           'PuL' = "posterior",
                           'PuM' = "posterior",
                           'VA' = "ventral",
                           'VAmc' = "ventral",
                           'VLa' = "ventral",
                           'VLp' = "ventral",
                           'VM' = "ventral",
                           'VPL' = "ventral",
                           'Whole_thalamus' = "Whole_Thalamus")) %>%
  unite("hemi_group", c(hemi, group), sep = "_", remove = FALSE, na.rm = FALSE) %>%
  unite("group_labeller", c(group, hemi), sep= " ", remove = FALSE, na.rm = FALSE)%>%
  mutate(region_type = dplyr::recode(region, 'AV' = "Subregion", 
                           'CL' = "Subregion",
                           'CM' = "Subregion",
                           'CeM' = "Subregion",
                           'L-Sg' = "Subregion",
                           'LD' = "Subregion",
                           'LGN' = "Subregion",
                           'LP' = "Subregion",
                           'MDl' = "Subregion",
                           'MDm' = "Subregion",
                           'MGN' = "Subregion",
                           'MV(Re)' = "Subregion",
                           'Pc' = "Subregion",
                           'Pf' = "Subregion",
                           'Pt' = "Subregion",
                           'PuA' = "Subregion",
                           'PuI' = "Subregion",
                           'PuL' = "Subregion",
                           'PuM' = "Subregion",
                           'VA' = "Subregion",
                           'VAmc' = "Subregion",
                           'VLa' = "Subregion",
                           'VLp' = "Subregion",
                           'VM' = "Subregion",
                           'VPL' = "Subregion",
                           'Whole_thalamus' = "Whole")) %>%
  dplyr::select(-c(region_recoded, analysis))
```

