---
title: "Thalamus Analysis"
author: "FG"
date: "5 3 2021"
output: 
  bookdown::html_document2:
    code_folding: hide
    toc: yes
    toc_depth: 5
    highlight: tango
    theme: paper
    fig_width: 20
    fig_height: 20
    fig_caption: yes
    number_sections: yes
    keep_md: yes
    df_print: default
---

```{r library, results='hide', message=FALSE, warning=FALSE} 
# library(devtools) # for development versions on GitHub
library(htmlTable)
#library(magick)
library(rmarkdown)
library(Hmisc)
# library(Rcmdr)
library(dplyr)
library(psych)
library(pastecs)
library(tidyr)
library(tidyverse)
library(ggplot2) # for data visualization (plotting)
library(car) # for statistical models (ANCOVA with type III errors), dummy coding
library(MatchIt) # for propensity (e.g. age, IQ) matching
# library(purrr)  # for applying functions and working with lists (e.g. map(), reduce(), transpose())
# library(stringr) # for string manipulation (e.g. str_detect(), str_extract(), str_subset(); string pattern - regular expressions)
# library(forcats) # for dealing with factors (e.g. reordering)
library(ggforce) # extends functionality of ggplot2 (e.g. spreads boxplot facets (facet_wrap) over multiple pages)
library(cowplot) # extends functionality of ggplot2 (e.g. create raincloud plots)
# library(PupillometryR) # for raincloud (flat violin) plots
library(ggsci) # alternative color tool/palettes for plotting (scientific journal style)
library(ggpubr) # extends functionality of ggplot2 (e.g. editing options for publication)
library(ggsignif) # for adding significance brackets and labels to ggplots (better compatible with geom_boxplot() than ggpubr)
# library(stargazer) # for creating aesthetic table outputs to summarize analyses
library(kableExtra) # for creating kables (customized tables in RMarkdown output with styling options)
library(bookdown) # support for numbering figure and table captions
library(rstatix) # for statistical tests (in combination with dplyr/piping and plotting) and summary statistics
# library(psych) # additional statistic options (e.g. correlation matrix, partial correlations)
# library(lme4) # for linear mixed modelling
library(MASS) # for robust linear model
library(broom) # tidy test output, e.g. of GLM
library(emmeans) # for estimated marginal means (EMM) (covariate-adjusted means in ANCOVA)
library(sfsmisc) # for p values of robust regression analyses
# library(broom.mixed) # tidy mixed model
library(scales) # for scale and percentage formatting (optional)
# library(corrplot) # generate correlation matrix/visualization
# library(ppcor) # for partial correlations
# library(effectsize) # for effect size statistics (e.g. eta squared, Cohen's d)
library(sjstats) # for effect size statistics
# library(pwr) # basic functions for power analysis
# library(broom)
# #library(matlib) # for Gram-Schmidt Orthogonalization of a Matrix
# library(tidystats) # make lists and dataframes out of statistical test results
# library(reshape2) # for transposing/melting data (can also be done with dplyr)
# library(RColorBrewer) # color tool/palettes for plotting
# library(captioner) # support for numbering figure and table captions (alternative to bookdown)
library(sqldf)
```


```{r}
QC_data <- get_redcap()%>%
  dplyr::select(point_of_research, redcap_event_name, participant_id, storage_name, mriqc_smri_thal_summary)%>%
  rename(thal_summary_redcap = mriqc_smri_thal_summary) %>%
  right_join(get_QC(), by =c("redcap_event_name", "participant_id", "storage_name"))%>%
  dplyr::select(-c(outlier_regions_symmetry, outlier_regions_volume))
  
  write_csv("data/QC_with_group.csv")
```
```{r}
a <- QC_data %>%
  filter(point_of_research %in% c("HCW","T1"))

b <- a %>%
  filter(is.outlier_cnr == FALSE & is.outlier_snr == FALSE)

c <- b %>% 
  filter(mriqc_smri_summary < 3)

d <- c %>%
  filter(n_outliers_symmetry_4.721sd < 1 & n_outliers_volume_4.721sd < 1) # only extreme volume & symmetry outliers!

d %>%
  filter(n_outliers_volume_2.698sd > 1 | n_outliers_symmetry_2.698sd > 1) %>%
  filter(mriqc_smri_thal_summary == 9)

d %>%
  filter(n_outliers_volume_2.698sd < 2 & n_outliers_symmetry_2.698sd < 2) %>%
  filter(mriqc_smri_thal_summary == 9)


```

*10. getting summary stats for thalamic nuclei*
```{r}
glmdata %>% 
  group_by(point_of_research, region, hemi_region, labeller) %>%
  rstatix::get_summary_stats(measure) %>%
  dplyr::select(hemi_region, labeller, point_of_research, n, mean, sd, min, max)
```


*11. select data for whole thalamus*
```{r}
whole.thal <- glmdata %>% filter(region == "Whole_thalamus")
```


# Descriptiv analysis for whole thalamus and subregions

1. Relationship of age and 
a) whole thal vol, 
b) thal subregions vol, 
c) eTIV

2. Frequency distributions of
a) whole thalamus volume
b) thalamic nuclei volume
c) eTIV

3. Q-Q Plots for
a) whole Thalamus
b) thalamic nuclei

## Impact of participant/patient age
### Relationship of age and whole Thalamus volume for acAN T1 and HCW
```{r}
whole.thal %>%
  ggplot(aes(x=age_at_date_of_research, y=measure, color=point_of_research)) +
  geom_point() + 
  geom_smooth(method=lm, aes(group=point_of_research, color=point_of_research, fill=point_of_research)) +
  scale_fill_jama(alpha=0.8) + 
  scale_color_jama(alpha=0.8) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep="~~~~")), label.y.npc = 1) +
  plot_theme_transparent + 
  labs(x = "Age at Date of Research", y = expression(paste("Whole thalamus volume in mm"^"3")))
ggsave(filename="figures/scatterplot_age_by_whole.png", width=20, height=20)
```


### Relationship of age and thalamic subregion volume for acAN T1 and HCW
```{r}
glmdata %>%
  ggplot(aes(x=age_at_date_of_research, y=measure, color=point_of_research)) +
  geom_point() + 
  geom_smooth(method=lm, aes(group=point_of_research, color=point_of_research, fill=point_of_research)) + 
  scale_fill_jama(alpha=0.8) + 
  scale_color_jama(alpha=0.8) +
  plot_theme_transparent + 
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep="~~~~")), label.y.npc = 1) +
  labs(x = "Age at Date of Research", y = expression(paste("Thalamus (sub-)region volume in mm"^"3"))) + 
  facet_wrap(~ region , scales="free")
ggsave(filename="figures/scatterplot_age_by_nuclei.png", width=20, height=20)
```


### Relationship of age and eTIV for acAN T1 and HCW
```{r}
whole.thal %>%
  ggplot(aes(x=age_at_date_of_research, y=e_tiv, color=point_of_research)) +
  geom_point() + 
  geom_smooth(method=lm, aes(group=point_of_research, color=point_of_research, fill=point_of_research)) +
  scale_fill_jama(alpha=0.8) + 
  scale_color_jama(alpha=0.8) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep="~~~~")), label.y.npc = 1) +
  plot_theme_transparent + 
  labs(x = "Age at Date of Research", y = expression(paste("eTIV in mm"^"3")))
ggsave(filename="figures/scatterplot_age_by_eTIV.png", width=20, height=20)
```

## Volume Distribution
### Whole Thalamus volume frequency distributions for acAN T1 and HCW 
```{r}
whole.thal %>%
  ggplot(aes(measure, fill = point_of_research)) + 
  geom_histogram(binwidth = 50, alpha=0.7) + 
  scale_fill_jama(alpha=0.8) + scale_color_jama(alpha=0.8) +
  plot_theme_transparent + 
  labs(x = expression(paste("Whole thalamus volume in mm"^"3")), y = "Frequency")
ggsave(filename="figures/boxplot_norm_whole_HC_T1.png", width=20, height=20)
```

### Whole Thalamus volume Frequency distributions with normal curve for HCW
```{r}
whole.thal %>%
  filter(point_of_research == "HCW") %>%
  ggplot(aes(x=measure, fill=point_of_research)) +
  geom_histogram(aes(y = ..density.. ), binwidth = 50,color = "grey40", fill ="lightblue4",size = 1) +
  scale_fill_jama() +
  plot_theme_transparent + 
  labs(x = expression(paste("HCW: whole thalamus volume in mm"^"3")), y = "Density") + 
  stat_function(fun = dnorm, args = list(mean = mean(whole.thal$measure), sd = sd(whole.thal$measure)), color = "grey20", size = 1)
ggsave(filename="figures/hist_norm_whole_T1.png", width=20, height=20)
```


### Whole Thalamus volume Frequency distributions with normal curve for acAN T1
```{r}
whole.thal %>%
  filter(point_of_research == "T1") %>%
  ggplot(aes(x=measure, fill=point_of_research)) +
  geom_histogram(aes(y = ..density.. ), binwidth = 50, color = "orange3", fill ="burlywood2",size = 1) +
  scale_fill_jama(alpha=0.8) + 
  scale_color_jama(alpha=0.8) +
  plot_theme_transparent + 
  labs(x = expression(paste("acAn T1: whole thalamus volume in mm"^"3")), y = "Density") + 
  stat_function(fun = dnorm, args = list(mean = mean(whole.thal$measure), sd = sd(whole.thal$measure)), color = "darkorange4", size = 1)
ggsave(filename="figures/hist_norm_whole_HCW.png", width=20, height=20)
```

### Whole Thalamus volume density plot for acAN T1 and HCW
```{r}
whole.thal %>%
  ggplot(aes(measure, color = point_of_research)) +
  geom_density() +
  scale_color_jama(alpha=0.8) +
  plot_theme_transparent
ggsave(filename="figures/whole_density.png", width=20, height=20)

```

### Thalamus (sub-)regions Histograms with normal curve
```{r figurehistothal, fig.cap="Histograms with normal curves for thalamus (nuclei) volumes", echo=TRUE, include=TRUE, message=FALSE, warning=FALSE}
histos <- glmdata  %>%
  group_by(region) %>% arrange(region, .by_group=TRUE) %>%
  mutate(point_of_research = dplyr::recode(point_of_research,
                                    "T1" = "AN",
                                    "HCW" = "HC"))

plotlist = list()

bins <- 30

for (Region in unique(histos$region)) {
  
  histdata <- histos %>% filter(region %in% Region)
  
  

  bw <- histdata %>% # individual binwidths per plot depending on the measure range
  summarize(bw = (max(measure) - min(measure))/bins) %>%
  pull(bw)
  
  mean_sd_T1 <- histdata %>% filter(point_of_research %in% "AN") %>%
    summarize(mean = mean(measure), sd = sd(measure))
  
  mean_sd_HC <- histdata %>% filter(point_of_research %in% "HC") %>%
    summarize(mean = mean(measure), sd = sd(measure))
  
  thalahist <- histdata %>%
    ggplot(aes(x=measure, fill=point_of_research)) +
    geom_histogram(aes(y=..density..), binwidth=bw, alpha=0.7, position="identity") +
    stat_function(color="grey35", size=1.5, fun=dnorm, args=list(mean=mean_sd_T1$mean, sd=mean_sd_T1$sd)) +
    stat_function(color="gray60", size=1.5, fun=dnorm, args=list(mean=mean_sd_HC$mean, sd=mean_sd_HC$sd)) +
    scale_fill_grey() +
    #scale_fill_jama() +
    scale_y_continuous(sec.axis = sec_axis(trans = ~. * bw )) +
    scale_x_continuous(breaks = extended_breaks(n=3)) +
    #scale_y_continuous(expand = c(0, 0)) +
    plot_theme_transparent3 +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank()) +
    theme(strip.text.x = element_text(size=30)) +
    theme(legend.position = "none") +
    #theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
    facet_wrap(~ labeller, scales="free_x", labeller = label_wrap_gen(width=15))
  
 plotlist[[Region]] = thalahist
 
}
histo.nuclei <- plotlist[1:25] 

ggarrange(plotlist=plotlist, common.legend = TRUE, legend="none", ncol=3, nrow=9) -> histoarrange
histoarrange <- annotate_figure(histoarrange, 
                bottom = text_grob(expression(paste("Volume in mm"^"3")), size=35),
                left = text_grob("Frequency density", size=35, rot=90),
                right = text_grob("Counts", size=35, rot=270))

ggsave(histoarrange, path="figures", filename="histograms_normal_nuclei+whole.png", width=35, height=45, dpi=200)

```

### Thalamus subregions volumes Bar charts for acAN T1 and HCW 
```{r}
glmdata%>%
  filter(region != "Whole_thalamus") %>%
  ggplot(aes(region, measure, fill = point_of_research)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") + 
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", position = position_dodge(width = 1), width = 0.5) +
  scale_fill_jama(alpha=0.8) + 
  scale_color_jama(alpha=0.8) +
  plot_theme_transparent + 
  labs(x = "Thalamus nuclei names", y = expression(paste("Thalamic nuclei volume in mm"^"3")))
ggsave(filename="figures/hist_AllNuclei.png", width=20, height=20)
```


### eTIV frequency distributions for acAN T1 and HCW (estimated Total Intracranial Volume, aka ICV)
```{r}
whole.thal %>%
  ggplot(aes(e_tiv, fill=point_of_research)) + 
  geom_histogram(binwidth = 10000) + 
  scale_fill_jama(alpha=0.8) + 
  scale_color_jama(alpha=0.8) +
  plot_theme_transparent + 
  labs(x = expression(paste("eTIV in mm"^"3")), y = "Frequency")
```


### Thalamus volume Q-Q Plots
*stands for Quantile - Quantile Plot. Plots the comulative values of our data against the cumulative probability of a particular distribution: in this case a normal distribution.*
*Data are ranked and sorted. Each value is compared to the expected value that the score should have in a normal distribution. Than plotted against each other.*
### Thalamus subregions volume Q-Q Plots
```{r}
glmdata %>% 
  mutate(point_of_research = fct_relevel(point_of_research, c("HCW", "T1"))) %>%
  ggplot(aes(sample=measure, group=point_of_research, color=point_of_research)) +
  stat_qq(geom="point", size=1.25, shape=16, alpha=1, position="identity") +
  stat_qq_line(size=1.25) +
  scale_fill_jama(alpha=0.8) + 
  scale_color_jama(alpha=0.8) +
  plot_theme_transparent + 
  facet_wrap(~ region , scales="free")

ggsave(filename="figures/nuclei_QQ_C.png", width=20, height=20)
```

